---
title: "Phyloseq"
author: "Angelica Torres"
date: "'r Sys.Date()'"
output: pdf_document
---

# Load required packages
```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("phyloseq")
```

# Load phyloseq
```{r}
library(phyloseq)
library(dplyr)
library(BiMiCo)
library(ggplot2)
```

# Load taxa and seqtab.nochim
```{r}
load("RData/taxa.RData")
load("RData/seqtab.nochim.RData")
```

# Import metadata
```{r}
metadata<-read.csv("metadata.csv", header=TRUE, row.names = 1)
```

# Create phyloseq object
```{r}
#make sure the seqtab.nochim and taxa objects are loaded
physeq <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows=FALSE), sample_data(metadata), tax_table(taxa))
physeq
```


# Remove the sequence itself and replace with ASV
```{r}
dna <- Biostrings::DNAStringSet(taxa_names(physeq))
names(dna) <- taxa_names(physeq)
physeq <- merge_phyloseq(physeq, dna)
taxa_names(physeq) <- paste0("ASV", seq(ntaxa(physeq)))
physeq
```

# Remove mitochondria and chloroplast matches.remove all non bacterial sequences
```{r}
physeq <- physeq %>% subset_taxa( Family!= "Mitochondria" | is.na(Family) & Order!="Chloroplast" | is.na(Order))
physeq
```

# Remove all non bacterial sequences
```{r}
physeq<-rm_nonbac(physeq)
physeq
```

# save physeq objects to load later
```{r}
save(physeq, file="RData/physeq.RData")
```

#load physeq objects to start here
```{r}
load("Rdata/physeq.RData")
```

# plot bar graph based on phylum
```{r}
plot_bar(physeq, fill = "Phylum") + geom_bar(aes(color=Phylum, fill=Phylum),stat="identity", position="fill")
```

# Create a barplot of relative abundance
```{r}
# convert to relative abundance
physeq_relabund <- transform_sample_counts(physeq, function(x) x/ sum(x))

# barplot
plot_bar(physeq_relabund, fill = "Phylum") + geom_bar(aes(color=Phylum,fill=Phylum), stat="identity", position="fill") + facet_wrap(~body.site, scales= "free")
```

# plot alpha diversity based on body.site
```{r}
plot_richness(physeq, x="body.site", color= "subject",
measures=c("Observed","Simpson", "Shannon"))
```

# plot alpha diversity based on subject
```{r}
plot_richness(physeq, x="subject", color= "body.site",
measures=c("Observed","Simpson", "Shannon"))
```

# test for normality
```{r}
alpha <- estimate_richness(physeq, measure=c("Observed","Simpson", "Shannon"))
```

# shapiro-wilk
```{r}
observed <- shapiro.test(alpha$Observed)
shannon <- shapiro.test(alpha$Shannon)
simpson <- shapiro.test(alpha$Simpson)

# print
print(observed)
print(shannon)
print(simpson)
```

#create data frames for statistical analysis
```{r}
# extract sample information from the physeq object
samples <- sample_data(physeq)

# if samples is a phyloseq sample_data object, convert it to a data frame
if (class(samples) == "sample_data"){samples <- data.frame(sample_data(samples))}

# add a column to alpha with sample names
alpha$sample <- rownames(alpha)

# merge alpha diversity data and sample data
alpha <- merge(alpha, samples, by = "sample")
```

# perform statistics based on subject
```{r}
# Perform t/wilcox tests for each biodiversity index
test_observed <- wilcox.test(Observed ~ subject, data = alpha)
test_simpson <- wilcox.test(Simpson ~ subject, data = alpha)
test_shannnon <- t.test(Shannon ~ subject, data = alpha)

# Printing the results
print(test_observed)
print(test_simpson)
print(test_shannnon)
```

# perform statistics based on reported antibiotic usage
```{r}
# Perform t/wilcox tests for each biodiversity index
test_observed <- wilcox.test(Observed ~ reported.antibiotic.usage, data = alpha)
test_simpson <- wilcox.test(Simpson ~ reported.antibiotic.usage, data = alpha)
test_shannnon <- t.test(Shannon ~ reported.antibiotic.usage, data = alpha)

# Printing the results
print(test_observed)
print(test_simpson)
print(test_shannnon)
```

# test for body site
```{r}
kruskal.test(Simpson ~ body.site, data=alpha)
pairwise.wilcox.test(alpha$Simpson, alpha$body.site, p.adjust.method="holm")
kruskal.test(Observed ~ body.site, data=alpha)
pairwise.wilcox.test(alpha$Observed, alpha$body.site, p.adjust.method="holm")
shannonanova <- aov(Shannon ~ body.site, data=alpha)
summary(shannonanova)
TukeyHSD(shannonanova)
```

# Remove taxa with relative abundance <0.005%
```{r}
minTotRelAbun = .00005
x = taxa_sums(physeq)
keepTaxa = (x / sum(x)) > minTotRelAbun
physeqprune = prune_taxa(keepTaxa, physeq)
physeqprune
```

# Number of shared ASVs body.site (found in 25% or more)
```{r}
bodysite=ps_venn(
  physeqprune,
  "body.site",
  fraction = .25,
  weight = FALSE,
  relative = TRUE,
  plot = TRUE
  )
bodysite
```

# Bray Curtis Calculation
```{r message=FALSE}
set.seed(777)
dist = phyloseq::distance(physeqprune, method="bray", weighted=TRUE)
ordination = ordinate(physeqprune, method="NMDS", distance=dist)
```

# Bray Curtis Subject Plot
```{r}
braysubject=plot_ordination(physeqprune, ordination, color="subject") + theme_classic() +
  theme(strip.background = element_blank()) + stat_ellipse(aes(group=subject))
braysubject
```

# Bray Curtis Body Site Plot
```{r}
braybodysite=plot_ordination(physeqprune, ordination, color="body.site") + theme_classic() +
  theme(strip.background = element_blank()) + stat_ellipse(aes(group=body.site))
braybodysite
```

# Bray Curits Antiobiotic Usage Plot
```{r}
braysubject=plot_ordination(physeqprune, ordination,
color="reported.antibiotic.usage") + 
  theme_classic() + 
  theme(strip.background = element_blank()) +  stat_ellipse(aes(group=subject))
braysubject
```

# Bray Curtis Subject Statistics
```{r}
adonis2(dist ~ sample_data(physeqprune)$subject)
```

# Bray Curtis Body Site
```{r}
adonis2(dist ~ sample_data(physeqprune)$body.site)
ps.disper<-betadisper(dist, sample_data(physeqprune)$body.site)
permutest(ps.disper, pair=TRUE)
```

# Bray Curtis Antibitic Usage
```{r}
adonis2(dist ~ sample_data(physeqprune)$reported.antibiotic.usage)
```